% ************************** Thesis Acknowledgements **************************

\begin{acknowledgements}

    I dedicate this thesis to my mother, who has never ceased supporting me since my birth and will continue to do so in the future, and to my father, who has always believed in me and encouraged me to challenge myself constantly.
    I also want to convey my deepest gratitude to my advisors, Romain Rouvoy and Lionel Seinturier, for allowing me to complete my Ph.D. study. Thank you for all the advice and discussions that helped me complete this project. Thank you for your unwavering support in helping me overcome the obstacles; without your help, I would never have overcome these challenges. I genuinely appreciate everything you've done to make this journey joyful for me.
    I also thank Davidson Consulting for providing me with the opportunity to work on this project by providing financial and technical support for this thesis. Thanks to David Olivier for his constant interest in my work.
    Also, I'd like to thank Pierre Boulet, Chantal Taconet, Olivier Barais, and Thomas Degeul, who are on my thesis committee, for their time and insightful feedback, which will help me keep this research going.

    I would also like to thank the Spirals team and its many talents who welcomed, guided me, and inspired me with their professional and personal achievements; special thanks to Zakaria Ournani for his constant emotional and technical support during this whole journey.

    Lastly, I would like to warmly thank my friends who helped, guided, and encouraged me during this adventure. You believed in me even when I doubted myself, and for this, I will be eternally grateful to you. A special mention goes to my dear friend Oudjedi Damerdji Soufiane, who inspired me with his endeavors during his fight, May your memory remain with us. And my dear girlfriend, who has always been there to support me.




\end{acknowledgements}
